---
title: "AudioFeaturesEDA"
author: "Mary Solomon"
date: "1/17/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(data.table)
library(tidyverse)
library(ggplot2)
library(ggthemes)

gghist <- function(var, var_name){
  ggplot(data = data, aes(var)) + 
  geom_histogram(fill = 'white', col = 'black') +
  xlab(var_name) +
  theme_minimal()
}
```

### Import Data 
```{r}
data <- fread("audiofeatures_clean.csv")
head(data, n=5)
```

Let's look at a summary
```{r}
summary(data)
```


release date distribution
```{r}
#gghist(data$release_date, 'release date')
```


popularity distribution
```{r}
ggplot(data = data, aes(popularity)) + 
  geom_histogram(fill = 'white', col = 'black') +
  theme_minimal()
gghist(data$popularity, 'popularity')
hist(data$popularity)
```



duration distribution
```{r}
gghist(data$duration_ms, 'duration (ms)')
gghist(log(data$duration_ms), 'duration (ms) and log transform')
```



acousticness distribution
```{r}
gghist(data$acousticness, 'acousticness')
```

danceability distribution
```{r}
gghist(data$danceability, 'danceability')
```

energy distribution
```{r}
gghist(data$energy, 'energy')
```

instrumentalness distribution
```{r}
gghist(data$instrumentalness, 'instrumentalness')
gghist(log(data$instrumentalness), 'instrumentalness (log transform)')
```


key distribution (categorical 1-12)
```{r}
gghist(data$key, 'musical key')
```


liveness distribution
```{r}
gghist(data$liveness, 'liveness')
```

loudness distribution
```{r}
gghist(data$loudness, 'loudness')
```

mode distribution (also categorical)
```{r}
gghist(data$mode, 'musical mode')
```


speechiness distribution
```{r}
gghist(data$speechiness, 'speechiness')
gghist(log(data$speechiness), 'speechiness (log transform)')
```


tempo distribution
```{r}
gghist(data$tempo, 'tempo')
```



time_signature distirbution (categorical...)
```{r}
gghist(data$time_signature, 'time signature')
```


valence distribution
```{r}
gghist(data$valence, 'valence')
```


