---
title: "Predict Popularity"
author: "Mary Solomon"
date: "2/11/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(QuantPsyc)#for multivariate normality function
library(data.table)
library(tidyverse)
library(ggplot2)
library(ggthemes)
library(stringr)
library(scales)
```

data :D
```{r}
data <- fread("kpopdata.csv")
# data <- mutate(data, ArtistType = as.factor(ArtistType),
#                ArtistGender = as.factor(ArtistGender),
#                ArtistGen = factor(ArtistGen),
#                #release_date = as.POSIXct.Date(release_date),
#                key = factor(key, levels = 0:11),
#                mode = as.factor(mode),
#                time_signature = factor(time_signature, levels = c(1,3,4,5)))
```


select just audio features
```{r}
kpop <- dplyr::select(data, popularity, duration, acousticness, danceability, energy, instrumentalness, key, loudness, mode, speechiness, time_signature, tempo, valence)
```

General Assumptions:
continuous response: popularity score ranging from 0 - 100.
mix of categorical and continuous response.
the distribution of the variables are not normal, we will check for normality of error terms where appropriate.

Goal: create model for predicting popularity scores


# Linear Model?

multivariate normality
```{r}
mult.norm(kpop)$mult.test
```


no or little multicollinearity

no autocorrelation 

no homoscedasticity.
